---
title: "Formatting Gene Data" 
author: "Jon Judd" 
output: pdf_document 
---

The goal of this document is to create a centralized file of all and recessive genes that includes the gene symbol, ensembl ID, S_het, pLot, mutational length.

I will then take this file and then use effect sizes from burden tests to make plots modeling selection

# Packages

```{r}
rm(list = ls())

library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(readxl)
library(purrr)
library(tibble)
library(stringr)
library(forcats)
library(httr)
library(jsonlite)
```

# Loading base files

Let's start by just loading the requied files I have. I am not going to include 
burden results yet since those are in separate Nikhil file's

```{r}
# Load my 3 gene lists
universe_genes = read_tsv("/oak/stanford/groups/pritch/users/jjudd5/recessive_genes/gene_lists/universe_genelist.tsv", col_names = F)
recessive_genes = read_tsv("/oak/stanford/groups/pritch/users/jjudd5/recessive_genes/gene_lists/recessive_genelist.tsv", col_names = F)
recessive_lethal_genes = read_excel("/oak/stanford/groups/pritch/users/jjudd5/recessive_genes/gene_lists/recessive_lethal_genelist.xlsx")

# These gene lists all have a singlular issue with GBA1 being labeled as GBA, I need to fix this
universe_genes = universe_genes %>% mutate(X1 = if_else(X1 == "GBA", "GBA1", X1))
recessive_genes = recessive_genes %>% mutate(X1 = if_else(X1 == "GBA", "GBA1", X1))
recessive_lethal_genes = recessive_lethal_genes%>% mutate(Gene = if_else(Gene == "GBA", "GBA1", Gene))

# Load the gene symbol to ensemble file
symbol_conversion = read_tsv("/oak/stanford/groups/pritch/users/jjudd5/recessive_genes/gene_lists/symbol_ensembl_conversion.txt")

# Now load the required data relating to gene selection
shets = read_tsv("/oak/stanford/groups/pritch/users/jjudd5/recessive_genes/selection_data/s_het_estimates.genebayes.tsv")
targetsizes_plof = read_csv("/oak/stanford/groups/pritch/users/jjudd5/recessive_genes/selection_data/mutational_target_sizes_005.csv")
```


# Refine Recessive Gene List
The current recessive genes list offered by the MacArthur lab that I'm using is a bit vague and doesn't match up with OMIM data exactly. For this reason, I am going to match the gene list to specific autosomal/x-linked recessive genes using the OMIM API

I basically need to go gene by gene to see if they have a autosomal recessive disease linked to them in OMIM and no dominant diseases

```{r}
# Let's define some things
# Base API and key
api_key = "tcjE1WBLQhOKGi9ePsmtjA"
base_url = "https://api.omim.org/api/geneMap/search"

# Now pull all the genes that need to be checked
rec_genes = recessive_genes$X1

# String to save data as well as genes to save genes without proper informaiton
macarthur_verified_genes = character()
macarthur_verified_traits = data.frame()

# Loop through increments of the entries
for (gene in rec_genes){

  #Define some entry parameters
  params = list(
    search = gene,
    apiKey = api_key,
    format = "json",
    start = "0",
    limit = "100"
  )
  
  # Get data
  response = GET(url = base_url, query = params)
  # Parse JSON content
  data <- fromJSON(content(response, "text"))
  #Get Gene specific data
  geneData = data$omim$searchResponse$geneMapList$geneMap
  
  # I want to go to the next gene if there is no gene or phenotype data
  if (is_empty(geneData)) {
    next
  }
  if (is_empty(geneData$phenotypeMapList)){
    next
  }
  
  # Extract relevant phenotype data
  phenotypes = geneData$phenotypeMapList %>% 
    map_dfr(., ~as_tibble(.x)) %>% 
    unnest(cols = everything())
  
  # Figure out the inheritance patterns
  inheritance = phenotypes$phenotypeInheritance
  
  # Add to external gene list if the gene leads to Autosomal recessive disease
  if ("Autosomal recessive" %in% inheritance && 
      !any(str_detect(inheritance, "dominant"), na.rm = T)) {
    
    # Save the gene name
    macarthur_verified_genes = str_c(c(macarthur_verified_genes, gene))
    
    # Save the gene with the phenotypes and add a phenotypic series column and just make it null if something doesn't exist
    if (!("phenotypicSeriesNumber" %in% colnames(phenotypes))) {
        phenotypes = cbind(phenotypes,
                           "phenotypicSeriesNumber" = NA)
    }
    # Add on the gene name so I know how they are matched up
    phenotypes = phenotypes %>% 
      cbind(Gene = gene, .)
    
    # Save the phenotypes
    macarthur_verified_traits = rbind(macarthur_verified_traits,
                               phenotypes)
  }
}

# Unique recessive genes
macarthur_verified_genes = macarthur_verified_genes %>% 
  unique() %>% 
  sort(.)

# Unique recessive phenotypes
macarthur_verified_traits = macarthur_verified_traits %>% 
  filter(!str_detect(phenotype, "\\?"),
         !str_detect(phenotype, "\\["),
         !str_detect(phenotype, "\\{")) %>% 
  unique(.) 

```


# Alternate Recessive Gene List
I'm increasingly doubting the original list of genes offered by the MacArthur gene list so I'm going to create a new list of recessive genes where I basically go through all the entries of OMIM and save all the genes that cause "Autosomal recessive" traits and no dominant traits
```{r}
# Let's define some things
# Base API and key
api_key = "tcjE1WBLQhOKGi9ePsmtjA"
base_url = "https://api.omim.org/api/geneMap/search"


# String to save data as well as genes to save genes without proper informaiton
omim_rec_genes = character()
omim_rec_phenotypes = data.frame()

# Loop through increments of the entries for all 18,577 OMIM entries
for (segment in seq(0, 18577 - 1, by = 100)) { 

  # Check where I am
  message("Current Segment:", segment)
  
  #Define some entry parameters
  params = list(
    search = "*",
    apiKey = api_key,
    format = "json",
    start = segment,
    limit = "100"
  )
  
  # Get data
  response = GET(url = base_url, query = params)
  # Parse JSON content
  data <- fromJSON(content(response, "text"))
  # Pull all gene data
  geneData = data$omim$searchResponse$geneMapList$geneMap
  
  # And then I want to search through each gene entry to see if they match my interested phenotype criteria
  for (gene in 1:nrow(geneData)){
  
    # Pull the current gene being searched
    curr_gene = geneData$approvedGeneSymbols[gene]
    
    # I want to go to the next gene if there is no phenotype data
    if (is.null(geneData$phenotypeMapList[[gene]])){
      next
    }
    
    # Extract relevant phenotype data
    phenotypes = geneData$phenotypeMapList[gene] %>% 
      map_dfr(., ~as_tibble(.x)) %>% 
      unnest(cols = everything())
    
    # Pull inheritance patterns
    inheritance = phenotypes$phenotypeInheritance
    
    # Add to external gene list if the gene leads to Autosomal recessive disease
    if ("Autosomal recessive" %in% inheritance && 
        !any(str_detect(inheritance, "dominant"), na.rm = T)) {
      
      # Save the gene name
      omim_rec_genes = str_c(c(omim_rec_genes, curr_gene))
      
      # Save the gene with the phenotypes and add a phenotypic series column and just make it null if something doesn't exist
      if (!("phenotypicSeriesNumber" %in% colnames(phenotypes))) {
          phenotypes = cbind(phenotypes,
                             "phenotypicSeriesNumber" = NA)
      }
      # Add on the gene name so I know how they are matched up
      phenotypes = phenotypes %>% 
        cbind(Gene = curr_gene, .)
      
      # Save the phenotypes
      omim_rec_phenotypes = rbind(omim_rec_phenotypes,
                                 phenotypes)
    }
  }
}

# Unique recessive genes
omim_rec_genes = omim_rec_genes %>% 
  unique() %>% 
  sort(.)

# Unique recessive phenotypes
omim_rec_phenotypes = omim_rec_phenotypes %>% 
  filter(!str_detect(phenotype, "\\?"),
         !str_detect(phenotype, "\\["),
         !str_detect(phenotype, "\\{")) %>% 
  unique(.) 


```


And then I want a quick section to compare the results from the two analyses looking at how to pick a gene list
```{r}
print(
  paste0("According to the MacArthur list of recessive genes verified through OMIM, there are: ",
      (macarthur_verified_genes %>% length(.)), " unique genes and ",
      (macarthur_verified_traits %>% pull(phenotypeMimNumber) %>% unique() %>% length()), " unique traits."
      )
)

print(
  paste0("According to a search through all of OMIM, there are: ",
      (omim_rec_genes %>% length(.)), " unique genes and ",
      (omim_rec_phenotypes %>% pull(phenotypeMimNumber) %>% unique() %>% length()), " unique traits."
      )
)
```



# Combine metrics

With everything loaded, I now have the opportunity to merge them together and then
note which genes in the universe of genes are recessive

The list of recessive genes are those that I just went and verified through OMIM

```{r}
# Quick formatting thing. I only need the gene symbols and ensembl id's from the conversion file
symbol_ensembl = symbol_conversion %>% select(`Approved symbol`, `Ensembl gene ID`)

# Start with the universe gene list, merge on the sybmol_conversion, and then add shets and target sizes
merged_dataset = universe_genes %>% 
  rename(symbol = X1) %>% 
  left_join(symbol_ensembl, by = c("symbol" = 'Approved symbol')) %>% 
  rename(ensg = 'Ensembl gene ID') %>% 
  left_join(shets, by = c('ensg' = 'ensg')) %>% 
  left_join(targetsizes_plof, by = c('ensg' = 'gene'))

# I need sqrt of the mutational target size so I will also include that
merged_dataset = merged_dataset %>% mutate(sqrt_mt = sqrt(mutational_target_size))

# And lastly, I want to also annotate which of these genes are part of the verified recessive gene list and if they are part of the recessive lethal genelist
merged_dataset = merged_dataset %>% 
  mutate(macarthur_rec = factor(ifelse(symbol %in% macarthur_verified_genes, 1, 0)),
         omim_rec = factor(ifelse(symbol %in% omim_rec_genes, 1, 0)),
         rec_lethal = factor(ifelse(symbol %in% recessive_lethal_genes$Gene, 1, 0)))
```

Quick testing plot code
```{r}
ggplot(merged_dataset) +
  geom_point(aes(x = sqrt_mt, y = aggregate_lof_frequency, colour = macarthur_rec)) + 
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
  theme_minimal()
```

# Output data

Let's save this dataset and then use it later on in case I want to generate new datasets
```{r}
write_tsv(merged_dataset, "/oak/stanford/groups/pritch/users/jjudd5/recessive_genes/gene_selection_datasets/gene_selection_data_2024.11.22.tsv")
```
